---
title: "rankrate: Joint Statistical Models for Preference Learning with Rankings and Ratings"
output: rmarkdown::html_vignette
bibliography: rankrate.bib
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

This package allows for joint modeling of ranking and rating preference data via the Mallows-Binomial model [@pearce2022unified]. Functions in the package may be used for density calculation, random data generation, and fitting the Mallows-Binomial model to data via multiple exact and approximate methods. Uncertainty quantification and estimation of confidence intervals is also possible via the nonparametric bootstrap, whose asymptotic validity was proven in @pearce2022validity. Additionally, the package includes 3 "toy" data sets and 1 real data from the American Institute of Biological Sciences, which were all studied in @gallo2022new.

For more details on how to use this package, see the [tutorial](https://pearce790.github.io/rankrate/articles/tutorial.html). 

A published version of the package may be install from CRAN, or a development version for the most up-to-date functionality:
```{r, eval=FALSE}
# CRAN version
install.packages("rankrate")  

# Development version
install.packages("devtools")  
devtools::install_github("pearce790/rankrate")
```

After installation, load the package with the following code:
```{r, out.width="90%", dpi=300}
library(rankrate)
```

# References
